### 2018-12-19 Thesis/Figures/Cavity.tex
Added laser cavity figure

\begin{tikzpicture}
% Two laser loops
\draw [rounded corners=4mm] (0,0) rectangle ++(9,4);
\draw [rounded corners=4mm] (0,0) rectangle ++(-1,4);

% Gain
\draw (0.5,2.25) circle (0.5cm);
\draw (0.5,2) circle (0.5cm) node [anchor=west,xshift=0.5cm,align=center] {Er Doped \\ Fibre};
\draw (0.5,1.75) circle (0.5cm);

% Modulator and pump
\filldraw[fill=white, draw=black] (3.5,-0.75) rectangle ++(2,1.5) node [midway] {Modulator};
\filldraw[fill=white, draw=black] (-1.6,1.5) rectangle ++(1.2,1) node [midway] {Pump};

% Coupler and output
\draw[-{Latex[length=2.5mm]}] (4.5,4) -- (4.5,5.5) node [pos=0.75,anchor=west,xshift=0.25cm] {Laser Output};
\draw[densely dashdotted] (4,3.5) -- (5,4.5) node [pos=1,anchor=north,yshift=-0.75cm,align=center] {Optical \\ Coupler};

% Circulator
\filldraw[fill=white, draw=black] (9,2) circle (0.5cm) node [anchor=west,xshift=0.5cm,align=center] {Optical \\ Circulator};
\draw (9,2.5) arc (0:-90:0.5cm);
\draw (9,1.5) arc (0:90:0.5cm);

% Grating
\draw (8.5,2) -- (6.5,2) node [pos=0.5,anchor=south,yshift=0.25cm,xshift=-0.15cm] {CFBG};
\foreach \i in {0,...,13}
  \draw (6.5 + \i*\i/100,1.75) -- (6.5 + \i*\i/100,2.25);

% Arrows
\draw[-{Latex[length=2.5mm]}] (2.16,0) -- (2.15,0);
\draw[-{Latex[length=2.5mm]}] (2.39,4) -- (2.4,4);
\draw[-{Latex[length=2.5mm]}] (6.76,0) -- (6.75,0);
\draw[-{Latex[length=2.5mm]}] (7.09,4) -- (7.1,4);
\draw[-{Latex[length=2.5mm]}] (-0.649,0) -- (-0.65,0);
\draw[-{Latex[length=2.5mm]}] (-0.351,4) -- (-0.35,4);
\draw[-{Latex[length=2.5mm]}] (8.807,2.11) -- (8.8,2.1);
\draw[-{Latex[length=2.5mm]}] (8.96,1.712) -- (8.97,1.7);
\end{tikzpicture}