### 2019-01-21 Thesis/Appendices/Dispersion.tex
Added appendices

% !TeX root = ../Thesis.tex

\chapter{Spread Due to Dispersion}

\cite{integrals}