### 2019-02-11 Thesis/Figures/Hologram.tex

% !TeX root = ../Thesis.tex

\begin{tikzpicture}

\draw[pattern=north east lines, pattern color=blue] (-4,0.5) rectangle (4,1); % Top Insulation
\draw[pattern=north west lines, pattern color=blue] (-4,-0.5) rectangle (4,-1); % Bottom Insulation
\draw[pattern=mydots, pattern color=OliveGreen] (-4,-0.5) rectangle (4,0.5); % Core

\foreach \i in {0,...,3}{
  % Rays
  \draw[->, >=stealth, magenta, thick] (-2.25 + \i,4.5) -- (-1.25 + \i,-1.5);
  \draw[->, >=stealth, magenta, thick] (-0.75 + \i,4.5) -- (-1.75 + \i,-1.5);
}
% Labels
\draw[->, >=stealth, thick] (3,-1.5) -- (2,0) node[pos=-0.2] {Ge Doped Core};
\draw[->, >=stealth, thick] (-3,-1.5) -- (-2,-0.75) node[pos=-0.3] {Cladding};
\draw[->, >=stealth, thick] (-3,2.5) -- (-1.82,2) node[pos=-0.4] {UV Light};
\end{tikzpicture}