### 2019-03-01 Thesis/Figures/Emission.tex
Added figure for spontaneous / stimulated emission

% !TeX root = ../Thesis.tex

\begin{tikzpicture}
% Energy levels
\draw [dashed] (0,0) -- (12,0) node [pos = -0.05] {$E_1$};
\draw [dashed] (0,4) -- (12,4) node [pos = -0.05] {$E_2$};

% Transition arrows
\draw[->, >=stealth] (2,0) -- (2,3.8);
\draw[<-, >=stealth] (6,0.2) -- (6,4);
\draw[<-, >=stealth] (10,0.2) -- (10,4);

% Circles
% Bottom
\filldraw[fill=white, draw=black, dashed] (2,0) circle (0.2);
\filldraw[fill=blue, draw=black] (6,0) circle (0.2);
\filldraw[fill=blue, draw=black] (10,0) circle (0.2);
% Top
\filldraw[fill=blue, draw=black] (2,4) circle (0.2);
\filldraw[fill=white, draw=black, dashed] (6,4) circle (0.2);
\filldraw[fill=white, draw=black, dashed] (10,4) circle (0.2);

% Photons
\draw [domain=0.5:2, samples=250, red, thick, ->, >=stealth] plot (\x, {-0.5*sin(4*pi*deg(\x-2))});
\draw [domain=6:7.5, samples=250, red, thick, ->, >=stealth] plot (\x, {4-0.5*sin(4*pi*deg(\x-6))});
\draw [domain=8.5:10, samples=250, red, thick, ->, >=stealth] plot (\x, {4-0.5*sin(4*pi*deg(\x-10))});
\draw [domain=10:11.5, samples=250, red, thick, ->, >=stealth] plot (\x, {4-sin(4*pi*deg(\x-10))});

% Labels
\node at (2, -1){Pumping};
\node [align=center] at (6, -1){Spontaneous \\ Emission};
\node [align=center] at (10, -1){Stimulated \\ Emission};
\end{tikzpicture}