### 2019-05-22 Thesis/Intro/Introduction.tex
Made some edits.

% !TeX root = ../Thesis.tex

\chapter{Introduction}
\begin{figure}[tbp]
\centering
\input{./Figures/Emission}
\caption[Comparison of pumping, spontaneous emission, and stimulated emission.]{In pumping, incident photons are absorbed by electrons which are then excited into a higher energy state (the electrons may also be excited by an electric field). In spontaneous emission, an electron in an excited state emits a photon and returns to the lower energy state. Finally, in stimulated emission, an incident photon interacts with an electron in an excited state, the electron donates its energy to the photon and returns to the lower energy state.}
\label{fig:emission}
\end{figure}

The word laser was originally an acronym for \emph{Light Amplification by Stimulated Emission of Radiation}. The cavity of a standard laser---such as a HeNe laser---is composed of four main components: a highly reflective mirror, a partially reflecting mirror, something to supply the pumping energy, and the gain medium. The two mirrors are at opposite ends of the cavity and serve to contain most of the light to the cavity while allowing a small amount to exit, this is the output of the laser. The pump energy is generally supplied by either an external light source, or an electric field. The gain medium is what amplifies the laser light and consists of gases, such as Helium and Neon, or a crystal, such as Neodymium-doped Yttrium Aluminium garnet (Nd:YAG). \\

More precisely, the atoms in the gain medium are pumped with energy raising the electrons to an excited state. For stimulated emission an incident photon triggers the emission of a photon from this excited electron \cite{alazzawi}. This new photon absorbs the electron's energy as it returns to its ground state, amplifying the light. Furthermore, the photon is released with the same direction and phase as the incident one, maintaining coherence \cite{alazzawi}. These processes are highlighted in Figure \ref{fig:emission}. The amplified light perpetually bounces between the two mirrors becoming more amplified with each interaction with an excited electron, a fraction of this light is able to pass through the partially reflecting mirror to become the output of the laser. \\

This is how laser light is generated, but, laser light has two fundamental characteristics that regular light does not. The first is that the light is highly monochromatic---the light contains (ideally) a single frequency. This is in contrast with an incandescent light bulb, for example, which emits light at all wavelengths with relative intensities given by Planck's law of black-body radiation. The other key feature is coherence, in which all the peaks and troughs of the light overlap giving very strong constructive interference---this is why laser light can be so intense. Something like a light emitting diode (LED) does not produce laser light because, while monochromatic, the light is not coherent. The monochromatic nature is a result of how the light is generated, the frequency of the light emitted directly corresponds to the energy difference between the excited and ground states of the gain medium (as with LEDs). Note that this correspondence means that  the frequency the laser operates at is fixed, and cannot be changed. Coherence is achieved in the laser cavity---the cavity is carefully designed so the light resonates with the length of the cavity. This causes all the peaks and troughs to align, and interfere constructively. \\

\section{Tuneable Lasers}
\begin{figure}[tbp]
\centering
\input{./Figures/Cavity}
\caption[Typical cavity of a tuneable laser.]{Typical cavity of a tuneable laser. The pulses travel clockwise around the loop.}
\label{fig:cavity}
\end{figure}

A tuneable laser on the other hand, has the ability to vary its wavelength by up to about 100 nanometres \cite{bohun, burgoyne2010, yamashita}. Tuneable lasers in fact lase at all of these different frequencies simultaneously. This tuneability is quite useful and has applications in spectroscopy and high resolution imaging such as coherent anti-Stokes Raman spectroscopy and optical coherence tomography \cite{bohun, burgoyne2014, yamashita}, as well as communications and diagnostics of ultra fast processes \cite{silfvast}. However, having the capacity to tune the wavelength comes with a cost, tuneable lasers vary considerably from standard lasers in both construction and operation, but, the central ideas are the same. A typical tuneable laser cavity is shown in Figure \ref{fig:cavity}. Instead of being perpetually reflected back and forth, the light travels in a circular fashion around the cavity, this is a so called ring laser. The gain medium is still excited by an external power source, and the output passes through a partially reflecting mirror to become the output. However, a tuneable laser also contains a few additional components, each of which will be described in the following subsections.

\subsection{Optical Coupler and Laser Output}
The optical coulper is a device that splits its input into two outputs---one continuing through the laser cavity, while the other exits the cavity to become the output of the laser. There are multiple devices that can accomplish this, however, the simplest is a partially reflecting mirror \cite{alazzawi}. These mirrors are characterized by their reflection coefficient, $R$. In the schematic shown in Figure \ref{fig:cavity}, the part of the signal that is reflected exits the cavity, whereas, the part of the signal transmitted remains within the cavity. \\

\subsection{Modulator}
The optical pulse in the cavity can be modulated with one of two methods---amplitude modulation, or phase modulation. Mathematically these can be considered equivalent since they are Fourier transforms of each other, but in practice these are accomplished by different apparatuses. For us, we shall assume the pulses are amplitude modulated. Generally, the pulse can be amplitude modulated using one of two techniques. The first is with an acousto-optic modulator which utilizes the acousto-optic effect where the index of refraction of the fibre is varied by a sound wave \cite{hausbook, karim}. The second---and more common---method uses the electro-optic effect and is called an electro-optic modulator. These work the same way, the fibre inside the modulator responds to an applied electric field, modifying the index of refraction \cite{agrawal2002, goldstein, hausbook, karim}. The shape of the electrical (sound) pulse can be controlled in such a way to provide the desired effect of modulation. \\

Moreover, a Mach--Zehnder interferometer can be used in tandem with an acousto- or electro-optic modulator to modulate an optical pulse. The incident pulse is separated with a Y-waveguide, one of the two branches is then modulated, and finally the two branches are recombined with another Y-waveguide \cite{alazzawi, hausbook, karim}. In the absence of a modulator on one of the branches the two pulses recombine into the original pulse. However, with modulation the two pulses are no longer identical, and so when the recombine, they interfere. This interference can be used to further modulate the pulse \cite{agrawal2002, agrawal2013, hausbook, karim}. \\

\subsection{Fibre Bragg Grating}
A fibre Bragg grating (FBG) is an optical fibre where the refractive index varies periodically along its length \cite{ferreira}. To achieve this, typically silica fibres are doped with Germanium, which when exposed to intense ultraviolet (UV) light alters the refractive index of the core \cite{becker, starodoumov}. The photosensitivity of the optical fibres can be increased by more than an order of magnitude by the Germanium doping \cite{becker, ferreira}. \\

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/PhaseMask}
\caption{\textbf{Phase-Mask:} The phase-mask refracts the UV light onto the core.}
\label{fig:phasemask}
\vspace{10mm}
\end{subfigure}
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Hologram}
\caption{\textbf{Holographic Side Exposure:} Two incident beams intersect at the core.}
\label{fig:holographic}
\end{subfigure}
\caption[Manufacture methods of fibre Bragg gratings.]{Depictions of the two most common methods for manufacturing FBGs. The UV light is focused on the core causing periodic constructive and destructive interference. The UV light alters the refractive index of the core, causing a sinusoidal variation along the length.}
\label{fig:fbgmake}
\end{figure}

FBGs are manufactured using one of two methods---the phase-mask method \cite{agrawal2002, alazzawi, becker, starodoumov}, or the holographic side exposure method \cite{agrawal2002, alazzawi, becker, ferreira, starodoumov}---these are shown in Figure \ref{fig:fbgmake}. Both methods cause the periodic nature of the refractive index through interference. In the phase-mask method (Figure \ref{fig:phasemask}) a single beam of UV light passes through the phase-mask which acts as a series of lenses, focusing the light at the core---this causes a sinusoidal interference pattern. Similarly, in the holographic side exposure method (Figure \ref{fig:holographic}), two beams of UV light are instead used to create the interference pattern. \\

The purpose of FBGs is that they act as reflective filters \cite{agrawal2002, alazzawi, ferreira, starodoumov}. Due to the periodicity of the refractive index, light with the corresponding wavelength will be reflected with all others passing through. This wavelength is defined by the Bragg condition \cite{agrawal2002, alazzawi, becker, ferreira, silfvast, starodoumov}:
\begin{align}
\label{eq:bragg}
\lambda_B = 2 \Lambda \bar{n},
\end{align}
where $\lambda_B$ is the Bragg wavelength, $\Lambda$ is the period of the grating, and $\bar{n}$ is the average index of refraction. This reflects wavelengths close to $\lambda_B$, known as the stop-band. \\

\subsubsection{Chirped Fibre Bragg Grating}
\begin{figure}[tbp]
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Unchirped}
\caption{Unchirped Gaussian pulse.}
\label{fig:unchirped}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Chirped}
\caption{Linearly chirped Gaussian pulse.}
\label{fig:chirped}
\end{subfigure}
\caption[Comparison of a chirped and unchirped Gaussian pulse.]{In a unchirped pulse the frequency is constant, however, in a chirped pulse the frequency varies along the envelope.}
\label{fig:chirp}
\end{figure}

Chirp is simply the term for a signal that has a non-constant frequency across it. Figure \ref{fig:chirp} shows examples of chirped and unchirped Gaussian pulses---the most common type of chirp is linear chirp, where the frequency varies linearly across the pulse. Because of this, the oscillations are characterized by $\exp \left( i C x^2 \right)$, where $C x$ is the linear variation of the wave number, and $C$ is the chirp parameter. By using a chirped phase-mask a chirped fibre Bragg grating (CFBG) can be created. Since the period of the refractive index varies across the CFBG, so does when the Bragg condition, \eqref{eq:bragg}, is satisfied. This causes most wavelengths to be reflected by a CFBG, but with each wavelength penetrating to a different depth. A consequence of this is that a time delay is created between wavelengths---this is depicted in Figure \ref{fig:cfbg} with the upper portion showing  the refractive index as a function of the depth. In this orientation, the red (dashed) wave is unable to penetrate as far as the blue (solid) wave since each wave is reflected where it matches the frequency of the refractive index.  \\

\begin{figure}[tbp]
\centering
\input{./Figures/CFBG}
\caption[Chirped fibre Bragg grating.]{Chirped fibre Bragg grating. Light is reflected when the wavelength coincides with the Bragg wavelength. This causes light with different wavelengths to penetrate to different depths. In turn, dispersion is either heightened or compensated---depending on the orientation.}
\label{fig:cfbg}
\end{figure}

Furthermore, the speed of light in an optical fibre is slightly dependent on the wavelength---this causes light with a longer wavelength to travel faster, and is known as chromatic dispersion. This is a large problem in fibre optic communications, the signal can spread and potentially becomes uninterpretable after large distances. However, chromatic dispersion can be counteracted using a CFBG \cite{agrawal2002, alazzawi, becker, starodoumov} (in the opposite orientation of Figure \ref{fig:cfbg}). By forcing the longer wavelengths to travel farther the dispersion can be reversed, restoring the original signal. In one experiment \cite{dong}, a signal was successfully transmitted over $109$ km at $40$ Gb/s by compensating the dispersion with two  $40$ cm CFBGs. Over this distance, the pulse would have spread to about $55$ times its original width, and could only have been transmitted $4$ km at that bit rate. At reduced bit rates however, a $10$ cm CFBG can compensate the dispersion of $300$ km of fibre \cite{agrawal2002}. However, in our case, we wish to accelerate the dispersion, simulating hundreds of metres of fibre, so the CFBG is used in the orientation shown in Figure \ref{fig:cfbg}. \\

\subsection{Optical Circulator}
\begin{wrapfigure}{O}{0.5\textwidth}
\centering
\input{./Figures/Circulator}
\caption{Symbol for a four port optical circulator.}
\label{fig:circulator}
\end{wrapfigure}
An optical circulator is a device that routes signals from port to port in a circular fashion \cite{agrawal2002, alazzawi, becker}, the symbol for a four port optical circulator is shown in Figure \ref{fig:circulator}. A signal entering from port 1 will be outputted from port 2; a signal entering from port 2 will exit from port 3; and so forth. Typically, optical circulators have three or four ports, with the first port being input only, and the final port being output only \cite{alazzawi}. Optical circulators are most commonly used with devices that reflect signals instead of transmit them. For example, a signal may enter through port 1, exit through port 2, be reflected by an FBG, re-enter port 2, and finally exit through port 3. \\

\subsection{Optical Amplifier and Pump Laser}
Optical amplifiers are of particular importance in fibre optic communications, they are used to restore the strength of a signal after it has been attenuated over large distances, or when a signal is divided into multiple paths \cite{alazzawi, starodoumov}. They are also more efficient, and introduce less noise than an electrical repeater. In this context, the optical amplifier provides the energy of the laser \cite{alazzawi}. Most commonly, optical amplifiers are created by doping a length of fibre (called the gain fibre) with a rare-earth element which receives power from a pump laser \cite{agrawal2002, alazzawi, starodoumov}. The most common dopant is Erbium, however, Ytterbium and Neodymium are also used---Holmium, Samarium, Thulium, and Tellurium are infrequently used as well \cite{agrawal2002}. \\

Erbium-doped fibre amplifiers (EDFAs) are used most widely since Erbium has a band gap that corresponds to $1.54$--$1.57$ $\mu$m, which is the preferred band for fibre optics since this has the least power loss \cite{agrawal2002, alazzawi, starodoumov}. The pump laser typically operates at $980$ nm or $1.48$ $\mu$m because these are able to transfer the most power (up to $100$ mW) into the fibre while introducing minimal noise \cite{agrawal2002, alazzawi, becker, starodoumov}. The pump power can be applied either forwards (with the laser), backwards (against the laser), or both \cite{alazzawi}, with each configuration having similar performance \cite{agrawal2002}. However, backwards pumping has slightly better performance at high powers when the gain begins to saturate\footnote{This concept will be discussed in Section \ref{chap:gain}.} \cite{agrawal2002}. Additionally, with the configuration shown in Figure \ref{fig:cavity}, the optical circulators can be used to isolate the pump circuit from the rest of the laser cavity. \\

\subsection{Mode-Locking}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML1}
\caption{$n = 4$}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\input{./Figures/ML2}
\caption{$n = 12$}
\end{subfigure}
\caption[Two examples of mode-locking.]{Two examples of mode-locking given by $\displaystyle \sum_{i = 1}^n \cos \left( 2 \pi i x \right)$.}
\label{fig:ml}
\end{figure}
One of the fundamental features of a laser is coherence. However, in the context of tuneable lasers it is perhaps less clear what is meant by coherence since the light is no longer monochromatic. Instead, we have a particular peak of all the frequencies aligned, that is, the phase shift between modes is zero. This process is called mode-locking---all of the different modes are locked together at a peak \cite{agrawal2013, hausbook, silfvast}. This of course creates a very intense and short pulse of light---as short as $5$ fs \cite{silfvast}. At this particular peak all of the frequencies constructively interfere as with a standard laser, however, deviations from the peak lead to destructive interference. Two examples of this are shown in Figure \ref{fig:ml}, and as expected, including more modes narrows and intensifies the pulse. \\