### 2019-05-22 Thesis/Thesis.tex
Added acknowledgements.

\documentclass[oneside]{UOIT_Thesis}

\usepackage{wrapfig}
\usepackage{lipsum}
\usepackage{amsthm}

% Load tikz and libraries
\usepackage{tikz}
\usetikzlibrary{patterns,decorations.pathreplacing}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{external}
\tikzexternalize[prefix=Figures/]

% New dots
\pgfdeclarepatternformonly{mydots}{\pgfqpoint{-1pt}{-1pt}}{\pgfqpoint{5pt}{5pt}}{\pgfqpoint{6pt}{6pt}}%
{
    \pgfpathcircle{\pgfqpoint{0pt}{0pt}}{.5pt}
    \pgfpathcircle{\pgfqpoint{3pt}{3pt}}{.5pt}
    \pgfusepath{fill}
}

% Code Stuff
\include{Listing}

\let\originalleft\left
\let\originalright\right
\renewcommand{\left}{\mathopen{}\mathclose\bgroup\originalleft}
\renewcommand{\right}{\aftergroup\egroup\originalright}

\providecommand{\df}{\textrm{d}}
\newcommand{\diff}[3][]{\frac{\textrm{d}^{#1}#2}{\textrm{d}{#3}^{#1}}}
\newcommand{\pdiff}[3][]{\frac{\partial^{#1}#2}{\partial{#3}^{#1}}}
\newcommand{\Es}{E_{\textrm{sat}}}
\newcommand{\FT}[1]{\mathcal{F}\left\{ #1 \right\}}
\newcommand{\FTi}[1]{\mathcal{F}^{-1}\left\{ #1 \right\}}
\newcommand{\Her}[2]{\widetilde{H}_{#1} \left( #2 \right)}
\newcommand{\eps}{\varepsilon}
\newcommand{\rect}[1]{\textrm{rect}\left( #1 \right)}

\providecommand{\bigO}[1]{\ensuremath{\mathop{}\mathopen{}\mathcal{O}\mathopen{}\left(#1\right)}}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\sech}{sech}

\newtheorem{theorem}{Theorem}[chapter]

\title{A New Method of Modelling Tuneable Lasers with Functional Composition}
\author{Brady Metherall}
\degree{Master of Science}
\faculty{Science}
\program{Modelling and Computational Science}

%\let\originalcite\cite
%\renewcommand{\cite}[1]{\mbox{\originalcite{#1}}}

\begin{document}

\frontmatter
\pagenumbering{roman} 
\maketitle

\makeabstract{
A new nonlinear model is proposed for tuneable lasers. Using the generalized nonlinear Schr\"{o}dinger equation as a starting point, expressions for the transformations undergone by the pulse are derived for each component within the laser cavity. These transformations are then composed to give the overall effect of one trip around the cavity. The linear version of this model is solved analytically, and the nonlinear version numerically. A consequence of the nonlinear nature of this model is that it is able to exhibit wave breaking which prior models could not. We highlight the rich structure of the boundary of stability for a particular plane of the parameter space.
}

\makeacknowledgements{
First and most importantly, I would like to thank my supervisor and mentor, Dr. Sean Bohun, for his assistance, guidance, and inspiration over the past two years. I would also like to thank my fellow modelling and computational science students for helping make the last two years enjoyable. \\

Lastly, I would like to thank my mom and dad, and grandparents for their endless support and encouragement over the past two years.
}

\makedeclaration

{\hypersetup{linkcolor=black}
\maketableofcontents
}

\mainmatter

\pagenumbering{arabic}
\doublespacing

\include{./Intro/Introduction}
\include{./Previous_Models/Previous}
\include{./Model/Model}
\include{./Nonlinear/Nonlinear}
\include{./Conclusion/Conclusion}

\phantomsection 
\addcontentsline{toc}{chapter}{References} 
\bibliography{Ref}

\appendix

\include{./Appendices/Lambert}
%\include{./Appendices/Dispersion}
%\include{./Appendices/Asymptotics}
\include{./Appendices/Hermite}

\chapter{Code}
\label{chap:code}
\lstinputlisting[language=Python]{../Lasers.py}

% Cite collections so they show up in the references
\nocite{abramowitz, gowers, hernandez}

\end{document}