### 2019-05-28 Thesis/Appendices/Asymptotics.tex
Finished corrections

% !TeX root = ../Thesis.tex
%\chapter{Asymptotic Expansions of the Variance}
%\label{chap:asymp}
We shall investigate the nature of the solution to \eqref{eq:var} both when $s \rightarrow 0$ (a very short CFBG), as well as when $s \rightarrow \infty$ (a very long CFBG). For ease of notation, \eqref{eq:var} is rewritten as
\begin{align}
\label{eq:asymp}
\frac{1}{4} x^4 = \eps \left( -x^3 + 5x^2 - 8x + 4 \right),
\end{align}
where $x = \sigma^2$, $\eps = s^4$, and $\eps \rightarrow 0$. Additionally, suppose $x$ can be expanded as a power series in $\eps$:
\begin{align}
\label{eq:series}
x = x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots
\end{align}
with $0 < \alpha < \beta$. Then, expanding \eqref{eq:asymp} using \eqref{eq:series} and collecting terms, at $\bigO{1}$,
\begin{align*}
\frac{1}{4} x_0^4 = 0,
\end{align*}
and so, $x_0 = 0$. Knowing this, we consider how to balance
\begin{align*}
\frac{1}{4} x_1^4 \eps^{4\alpha} = \eps \left( -x_1^3 \eps^{3\alpha} + 5x_1^2 \eps^{2\alpha} - 8x_1 \eps^\alpha + 4 \right).
\end{align*}
In order to have dominant balance it must be that the left hand side balances with the final term of the right hand side, that is, $4\alpha = 1$, or $\alpha = \frac{1}{4}$. Therefore, at $\bigO{\eps}$ $x_1^4 = 16$, or
\begin{align*}
(x_1 - 2)(x_1 + 2)(x_1^2 + 4) &= 0,
\end{align*}
and so, $x_1 = \pm 2, \, \pm 2 i$. However, we wish $x$ to be positive and real---recall that $x = \sigma^2$---thus, we take $x_1 = 2$. Now, the next lowest order of the left hand side must be $\eps^{3\alpha + \beta} = \eps^{3/4 + \beta}$. As with the previous iteration, this must balance with the lowest order term of the $8x$. Hence, $\frac{3}{4} + \beta = 1 + \alpha = \frac{5}{4}$, thus, $\beta = \frac{1}{2}$. Now, at $\bigO{\eps^{5/4}}$,
\begin{align*}
\frac{1}{4} 4 x_1^3 x_2 &= -8x_1,
\end{align*}
and with the choice of $x_1$, $x_2 = 2$. Combining these results gives $x = 2 \eps^{1/4} - 2 \eps^{1/2} + \bigO{\eps^{3/4}}$, or $\sigma^2 = 2 s \left( 1 - s \right) + \bigO{s^3}$,
this is shown in Figure \ref{fig:lims0}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_0}
\caption{Asymptotic expansion of the variance as $s \rightarrow 0$.}
\label{fig:lims0}
\end{figure}

We shall now consider the other limit, when $s \rightarrow \infty$. Using a similar substitution we instead write \eqref{eq:var} as
\begin{align}
\label{eq:asymp2}
\frac{1}{4} \eps x^4 = -x^3 + 5x^2 - 8x + 4,
\end{align}
where $x = \sigma^2$, and $\eps = s^{-4}$ instead so that we still have $\eps \rightarrow 0$. Furthermore, we must make a correction to the series expansion:
\begin{align}
\label{eq:series2}
x = \eps^{-\xi} \left( x_0 + x_1 \eps^\alpha + x_2 \eps^\beta + \cdots \right).
\end{align}
The reason for this is because the equation is now singular---when $\eps=0$ the equation transforms from a quartic to a cubic, losing a root. As with before, we obtain dominant balance when the quartic term balances with the cubic term which occurs when $1 - 4 \xi = -3 \xi$, or $\xi = 1$. Then to $\bigO{\eps^{-3}}$, expression \eqref{eq:asymp2} gives
\begin{align*}
x_0^3 (x_0 + 4) &= 0,
\end{align*}
we shall choose $x_0 = 0$---since $x \geq 0$. Continuing, we obtain a dominant balance when $\alpha = 1$. So the next condition occurs at $\bigO{1}$ where
\begin{align*}
0 = -x_1^3 + 5x_1^2 - 8x_1 + 4 = -(x_1 - 1)(x_1 - 2)^2,
\end{align*}
so either $x_1 = 1$ or $x_1 = 2$, we shall see in the next step that we can eliminate one of these. The next term obtains dominant balance when the left hand side balances with the linear term of the right hand side, that is, when $\beta = 2$. Now at $\bigO{\eps}$,
\begin{align*}
\frac{1}{4} x_1^4 = -3x_1^2 x_2 + 5\cdot2 x_1 x_2 - 8x_2 = -x_2 (3x_1 - 4)(x_1 - 2),
\end{align*}
where again, the coefficients come from the binomial expansions. From this it is clear that $x_2$ is unique provided $x_1 \ne 2$ and
\begin{align*}
x_2 &= \frac{-x_1^4}{4(3x_1 - 4)(x_1 - 2)}.
\end{align*}
Thus, $x_1 = 1$, $x_2 = -\frac{1}{4}$, and
\begin{align*}
x = 1 - \frac{1}{4} \eps + x_2 \eps^{\gamma - 1} + \cdots.
\end{align*}
Expanding one final term we find $\gamma = 3$, and to $\bigO{\eps^2}$,
\begin{align*}
\frac{1}{4} 4x_2 x_1^3 &= -(3x_3 x_1^2 + 3x_2^2 x_1) + 5(2x_3 x_1 + x_2^2) - 8x_3.
\end{align*}
Knowing the values of $x_1$, and $x_2$ this is simply an arithmetical calculation yielding $x_3 = \frac{3}{8}$. Finally,
\begin{align*}
x &= 1 - \frac{1}{4}\eps + \frac{3}{8}\eps^2 + \bigO{\eps^3},
\end{align*}
or,
\begin{align*}
\sigma^2 &= 1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}}
\end{align*}
this approximation is shown in Figure \ref{fig:limsinfty}. \\
\begin{figure}[tbp]
\input{./Figures/Lim_s_Infty}
\caption{Asymptotic expansion of the variance as $s \rightarrow \infty$.}
\label{fig:limsinfty}
\end{figure}

In conclusion, the asymptotic expansions of the observables are
\begin{align*}
\sigma^2 &=
\begin{cases*}
2s(1 - s) + \bigO{s^3} & $s \rightarrow 0$ \\
1 - \frac{1}{4s^4} + \frac{3}{8s^8} + \bigO{s^{-12}} & $s \rightarrow \infty$
\end{cases*}
& \phi &=
\begin{cases*}
\frac{1}{2}s - \frac{1}{24}s^3 + \bigO{s^5} & $s \rightarrow 0$ \\
\frac{\pi}{4} - \frac{1}{2s^2} + \frac{5}{12s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*} \\
C &=
\begin{cases*}
1 - s + \frac{1}{2}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{2s^2} - \frac{3}{8s^6} + \bigO{s^{-10}} & $s \rightarrow \infty$
\end{cases*}
& \zeta &=
\begin{cases*}
1 - \frac{1}{2}s + \frac{1}{8}s^2 + \bigO{s^3} & $s \rightarrow 0$ \\
\frac{1}{4^{1/4}s} - \frac{3 \cdot 4^{3/4}}{32 s^5} + \bigO{s^{-9}} & $s \rightarrow \infty$.
\end{cases*}
\end{align*}
The general form of these is as expected. When $s \rightarrow 0$, we effectively remove the CFBG from the cavity. This causes the pulse to be perpetually modulated, and so, the pulse approaches a $\delta$ function, and therefore, $\sigma^2 \rightarrow 0$. Moreover, since $\zeta$ is the ratio of the amplitude of the pulse before and after dispersion, $\zeta \rightarrow 1$. Similarly, $\phi \rightarrow 0$ since nothing in the circuit has the capacity to provide a phase shift. In the opposite extreme, $s \rightarrow \infty$, we have an incredibly long / powerful CFBG---this causes the pulse to broaden greatly, while conserving its energy. Because of this, the variance approaches unity because the shape is entirely determined by the modulation function. Additionally, $\zeta \rightarrow 0$ because the pulse becomes infinitely wide, and thus, the amplitude becomes infinitesimal.