### 2021-07-24 Papers/Journal/Proof_Errors_v2.tex
Added second iteration of proof revisions

\documentclass[11pt,a4paper]{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
%\usepackage[subfolder,cleanup]{gnuplottex}
%\usepackage{amsthm}
%\usepackage{enumitem}
%\usepackage{wrapfig}
%\usepackage{subcaption}
%\usepackage{hyperref}
%\usepackage{tikz}

% Nicer brackets for operators
\let\originalleft\left
\let\originalright\right
\renewcommand{\left}{\mathopen{}\mathclose\bgroup\originalleft}
\renewcommand{\right}{\aftergroup\egroup\originalright}

\newcommand{\Es}{E_{\textrm{sat}}} % Saturation energy

% Math operators
\providecommand{\bigO}[1]{\ensuremath{\mathop{}\mathopen{}\mathcal{O}\mathopen{}\left(#1\right)}}

% Macros
\newcommand{\diff}[3][\hspace{-0.5pt}]{\frac{\textrm{d}^{#1}#2}{\textrm{d}{#3}^{#1}}}
\newcommand{\pdiff}[3][\hspace{-0.5pt}]{\frac{\partial^{#1}#2}{\partial{#3}^{#1}}}
\newcommand{\df}{\, \textrm{d}}
%\newcommand{\eps}{\varepsilon}

% Row colouring in tables
%\usepackage[table]{xcolor}
%\rowcolors{2}{gray!25}{white}

% Margin size
\newgeometry{margin=2cm}

% Reference style
%\bibliographystyle{ieeetr}

\title{Predicting instabilities of a tunable ring laser with \\
an iterative map model \\ \Large Requested Revisions v2}
\author{Brady Metherall \and C. Sean Bohun}
\date{\today}

\begin{document}
\maketitle
\begin{itemize}
	\item Align $b=$ with $s=$ in equation 25.
	\item The second lines of equations 27 and 35 should be further right, for example:
	\begin{equation}
		\begin{gathered}
		\sigma^2 = \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} \\
		\hspace{15mm} - s^4 - s^2 \sqrt{4 + s^4}.
		\end{gathered}
	\end{equation}
	\item The two lines of equation 28 should be centred:
	\begin{equation}
		\begin{gathered}
			\widetilde{g}(\widetilde{A}) = \left( 1 + \frac{P_0 T_M}{\Es} \int_{-\infty}^\infty |\widetilde{A}|^2 \, \df \widetilde{T} \right)^{-1}, \\
			L_\textrm{disp} = \frac{T_M^2}{|\beta_2|}, \quad L_\textrm{NL} = \frac{1}{\gamma P_0}.
		\end{gathered}
	\end{equation}
	\item The font size in equations 27 and 28 seems to have been reduced, please return to the normal~size.
	\item Align figures 7 and 9.
	\item Most of the $\exp$, $\log$, $W$, and $\mathcal{O}$ seem to have additional space before the parenthesis, please remove the extra space. For example, compare the $W \left( a E_g \textrm{e}^{E_g} \right)$ terms in equations 33 and 38.
\end{itemize}

\noindent Please correct the size of the parentheses for:
\begin{itemize}
	\item Exponential on second line of equation 24.
	\item Logarithm in equation 37.
	\item $\exp \left( g_0 L_g \right)$ in equation 9, 13, 25.
	\item $\mathcal{O}$ in paragraph above equation 29 (two instances), and in Table 2.
\end{itemize}
I assume the additional space after the $\exp$ and such is a consequence of increasing the size of the parentheses. Please ensure the above changes do not add additional spacing.

\end{document}