### 2022-12-20 Thesis/Figures/Laser.tex
Missed this file somehow

\begin{tikzpicture}
% Light
\fill[red, path fading=fade left] (-3,-0.75/2.0) rectangle (3,0.75/2.0); % Gain
\fill[red, fill opacity=0.25] (-4,-0.75/2.0) rectangle (-3,0.75/2.0); % Left
\fill[red] (3,-0.75/2.0) rectangle (4,0.75/2.0); % Right
\fill[red] (4,-0.2) rectangle (6,0.2); % Right

\draw[fill=Goldenrod] (-4,1) rectangle (-3.8,-1); % Left
\fill[pattern=north west lines, pattern color=Goldenrod] (3.8,-0.2) rectangle (4,0.2); % Right Mid
\draw[fill=Goldenrod] (3.8,0.2) rectangle (4,1); % Right Top
\draw[fill=Goldenrod] (3.8,-0.2) rectangle (4,-1); % Right Bottom
\draw[pattern=mydots, pattern color=Purple] (-3,-0.75) rectangle (3,0.75); % Gain
\draw (-4,1) -- (4,1); % Top
\draw (-4,-1) -- (4,-1); % Bottom

% Labels
\draw[->, >=stealth, thick] (3.5,-1.5) -- (3.9,-0.75) node[pos=-0.3] {Partially Reflecting Mirror};
\draw[->, >=stealth, thick] (-3.5,-1.5) -- (-3.9,-0.75) node[pos=-0.3] {Mirror};
\draw[->, >=stealth, thick] (-1,-1.43) -- (-1,-0.5) node[pos=-0.3] {Gain Medium};
\draw[->, >=stealth, thick] (5,1.2) -- (5.5,0) node[pos=-0.3] {Laser Output};
\draw[->, >=stealth, thick] (0,1.2) -- (-3.5,0);
\draw[->, >=stealth, thick] (0,1.2) -- (3.5,0);
\draw (0,1.5) node {Laser Light};
\end{tikzpicture}